---
title: "data_extraction"
format: html
editor: visual
---

## Dependency Management

```{r}

renv::restore()

#list of used packages
library(jsonlite)
library(markdown)
library(dplyr)
library(purrr)
library(stringr)
library(tidyr)
library(irr)
library(xlsx)
```

## Column identifications

```{r}

# Column to connect papers on
id_col       <- "paper_id"
# metadata that will be excluded in the comparison
meta_exclude <- c("paper_title", "meta_id", "meta_year", "meta_authors",
                   "meta_title", "meta_journal", "meta_doi", "meta_reproducer")
compare_exclude <- c("coder_id", "meta_date", "_complete")

# Toggle for the pilot numbering fix (q3->q4, q4->q5, q5->q6, q6->q7).
# Set to FALSE for batches whose raw files already use the corrected numbering.
apply_pilot_rename <- TRUE

# Declarative recode spec, applied to each coder file's raw columns before
# suffixing. Add/remove entries here for future batches without touching
# load_coder_data itself.
# - type = "onehot": several 0/1 columns -> single column holding the mapped
#   code of whichever source column is 1; source columns are dropped.
# - type = "value": single text column -> its string values remapped to the
#   given codes; source column is dropped.
# Column names below are POST pilot-rename (see apply_pilot_rename above),
# since the rename runs before the recode inside load_coder_data.
recode_spec <- list(
  q1_supp       = list(type = "onehot", map = c(q1_yes = 1, q1_no = 0)),
  q2_code       = list(type = "onehot", map = c(q2_yes = 1, q2_no = 0)),
  q3_rm         = list(type = "onehot", map = c(q_readme_yes = 1, q_readme_no = 0, q_readme_unsure = 2)),
  q4_sufficient = list(type = "onehot", map = c(q4_ps = 1, q4_ci = 0, q4_un = 2)),
  q_itr_main    = list(type = "onehot", map = c(q_itr_main_yes = 1, q_itr_main_no = 0)),
  q_itr_supp    = list(type = "onehot", map = c(q_itr_supp_yes = 1, q_itr_supp_no = 0)),
  q5_source     = list(type = "value", col = "q5_source_0",
                        map = c("External Repository" = 1, "Journal website" = 2))
)

```

## Extraction and comparison functions

Don't change this part for new batches these are background functions.

<!--# Maybe move these functions to a separate script and load them in this script so new comparisons are easy to make -->

<!--# For final data the renumbering from this function should be removed -->

<!--# For final extraction: rename the variables to something that is easily recocnizable also for the other coders so discrepancy review is easier -->

```{r}

# Shift raw column names by the pilot numbering fix (q3->q4, q4->q5, q5->q6, q6->q7)
shift_pilot_numbering <- function(names_vec) {
  names_vec %>%
    str_replace("q6", "q7") %>%
    str_replace("q5", "q6") %>%
    str_replace("q4", "q5") %>%
    str_replace("q3", "q4")
}

# Collapse a set of 0/1 dummy columns into a single coded column
recode_onehot <- function(df, out_name, map) {
  present <- intersect(names(map), names(df))
  if (length(present) == 0) return(df)

  df[[out_name]] <- apply(df[present], 1, function(row) {
    hit <- present[which(row == 1)]
    if (length(hit) != 1) return(NA_real_)
    unname(map[hit])
  })
  df %>% select(-all_of(present))
}

# Remap a single text column's values to codes
recode_value <- function(df, out_name, col, map) {
  if (!col %in% names(df)) return(df)
  df[[out_name]] <- unname(map[df[[col]]])
  df %>% select(-all_of(col))
}

# Apply every entry in recode_spec to a coder's raw data
apply_recode_spec <- function(df, spec) {
  for (out_name in names(spec)) {
    entry <- spec[[out_name]]
    df <- switch(entry$type,
      onehot = recode_onehot(df, out_name, entry$map),
      value  = recode_value(df, out_name, entry$col, entry$map),
      stop("Unknown recode type: ", entry$type)
    )
  }
  df
}

# load, pilot-rename, recode, and suffix coder files
load_coder_data <- function(file, id_col, exclude, recode = TRUE, pilot_rename = apply_pilot_rename) {
  coder_label <- basename(file) %>% str_extract("(?i)coder\\d+") %>% str_to_lower()

  df <- fromJSON(file)$records %>%
    select(-any_of(exclude)) %>%
    # "not filled out" is stored as "" in the raw export, not JSON null;
    # convert to real NA so it's excluded from agreement/IRR, not compared as a value
    mutate(across(where(is.character), ~na_if(., "")))

  if (pilot_rename) {
    df <- df %>% rename_with(shift_pilot_numbering, -all_of(id_col))
  }
  if (recode) {
    df <- df %>% apply_recode_spec(recode_spec)
  }

  df %>% rename_with(~paste0(., "_", coder_label), -all_of(id_col))
}

# Combine coder files
combine_coders <- function(files, id_col, exclude = meta_exclude, recode = TRUE, pilot_rename = apply_pilot_rename) {
  map(files, load_coder_data, id_col = id_col, exclude = exclude,
      recode = recode, pilot_rename = pilot_rename) %>%
    reduce(full_join, by = id_col)
}

# Function to check if agreement should be checked and for which coders (only those that coded the paper) 
# This function is used in the next function
check_agreement <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) < 2) return(NA)   # 0 or 1 coder rated this — can't assess agreement
  length(unique(x)) == 1
}

# Add agreement column per variable
add_agreement_cols <- function(df, id_col, exclude = compare_exclude) {
  value_cols <- setdiff(names(df), id_col)
  stems <- value_cols %>% str_remove("_coder\\d+$") %>% unique() %>% setdiff(exclude)

  for (v in stems) {
    cols <- value_cols[str_detect(value_cols, paste0("^", v, "_coder\\d+$"))]
    if (length(cols) > 1) {
      df[[paste0(v, "_agreement")]] <- apply(df[cols], 1, function(x) check_agreement(x))
    }
  }
  df
}

# Reorder: id, then var_c1/c2/.../agreement per variable
reorder_combined <- function(df, id_col) {
  agreement_cols <- names(df)[str_detect(names(df), "_agreement$")]
  stems <- str_remove(agreement_cols, "_agreement$")

  ordered <- c(id_col, unlist(lapply(stems, function(v) {
    coder_cols <- sort(names(df)[str_detect(names(df), paste0("^", v, "_coder\\d+$"))])
    c(coder_cols, paste0(v, "_agreement"))
  })))
  df %>% select(all_of(ordered))
}

```

## Using the functions for a batch

<!--# Other version here that automatically lists the .json files in the folder instead of having to type them out? -->

```{r}

files_batch1 <- c("backup_CODER01_2026-08-06.json",
                   "backup_Coder03_2026-07-30.json",
                   "backup_Coder04_2026-07-27.json",
                  "backup_Coder02_2026-09-10.json")

combined <- combine_coders(files_batch1, id_col) %>%
  add_agreement_cols(id_col) %>%
  reorder_combined(id_col)
```

## Pilot numbering fix and recoding

The pilot renumbering (q3-\>q4, q4-\>q5, q5-\>q6, q6-\>q7) and the dummy/text recoding (`q1_supp`, `q2_code`, `q3_rm`, `q4_sufficient`, `q_itr_main`, `q_itr_supp`, `q5_source`) now happen inside `load_coder_data`, before the `_coderN` suffix is added — see `apply_pilot_rename` and `recode_spec` under Column identifications. Set `apply_pilot_rename <- FALSE` there, or pass `recode = FALSE` to `combine_coders()`, to turn either off for a batch that doesn't need it.

## Long-format data for review and IRR

```{r}

long_df <- combined %>%
  # combined mixes numeric/logical/free-text columns; pivot_longer needs one
  # common type for the unified value column, so coerce to character first
  mutate(across(-all_of(id_col), as.character)) %>%
  pivot_longer(
    cols = -all_of(id_col),
    names_to = c("variable", "field"),
    names_pattern = "^(.*)_(coder\\d+|agreement)$",
    values_to = "value"
  ) %>%
  pivot_wider(names_from = field, values_from = value)

```

## Review table of disagreements

```{r}

review_table <- long_df %>%
  filter(agreement == FALSE) %>%
  mutate(resolution = NA_character_, notes = NA_character_) %>%
  arrange(variable, .data[[id_col]])

by_variable <- review_table %>% count(variable, name = "n_discrepancies") %>% arrange(desc(n_discrepancies))
by_case     <- review_table %>% count(.data[[id_col]], name = "n_discrepancies") %>% arrange(desc(n_discrepancies))
remaining   <- sum(is.na(review_table$resolution))

# Save as an excel file
#write.xlsx(review_table, "discrepancies.xlsx")

```

## Inter-rater reliability (Krippendorff's alpha)

```{r}

irr_long <- long_df %>%
  select(-agreement) %>%
  pivot_longer(cols = starts_with("coder"), names_to = "coder", values_to = "value") %>%
  filter(!is.na(value))

compute_agreement_stats <- function(irr_long, id_col) {
  irr_long %>%
    group_by(variable) %>%
    group_modify(~ {
      mat <- .x %>%
        pivot_wider(id_cols = all_of(id_col), names_from = coder, values_from = value) %>%
        select(-all_of(id_col)) %>%
        as.matrix() %>% t()

      # fewer than 2 raters (common for sparsely-filled free-text fields) ->
      # agreement is undefined, not computable
      if (nrow(mat) < 2 || ncol(mat) < 1) {
        return(tibble(alpha = NA_real_, n_cases = ncol(mat), n_raters = nrow(mat)))
      }
      a <- irr::kripp.alpha(mat, method = "nominal")
      tibble(alpha = a$value, n_cases = a$subjects, n_raters = a$raters)
    })
}

agreement_stats <- compute_agreement_stats(irr_long, id_col) %>%
  left_join(by_variable, by = "variable") %>%
  rename(n_disagreements = n_discrepancies) %>%
  mutate(n_disagreements = replace_na(n_disagreements, 0))

```

## ------OLD scraps——

## Preparing dataframe for inserting coder data

```{r}

# Extracting variable names from json file

data_coder01 <- fromJSON("backup_CODER01_2026-08-06.json") %>% .$records %>%
  select(-c(paper_title, meta_id, meta_year, meta_authors, meta_title, meta_journal, meta_doi, meta_reproducer))
data_coder03 <- fromJSON("backup_Coder03_2026-07-30.json") %>% .$records %>%
  select(-c(paper_title, meta_id, meta_year, meta_authors, meta_title, meta_journal, meta_doi, meta_reproducer))
data_coder04 <- fromJSON("backup_Coder04_2026-07-27.json") %>% .$records %>%
  select(-c(paper_title, meta_id, meta_year, meta_authors, meta_title, meta_journal, meta_doi, meta_reproducer))

```

## Rename variables

```{r}

# suffix each coder's columns
c1 <- data_coder01 %>% rename_with(~paste0(., "_c1"), -paper_id)
c3 <- data_coder03 %>% rename_with(~paste0(., "_c3"), -paper_id)
c4 <- data_coder04 %>% rename_with(~paste0(., "_c4"), -paper_id)

# Combine into one dataframe
combined <- list(c1, c3, c4) %>% reduce(full_join, by = "paper_id")

# Variables to compare
var_compare <- c("q4_source_0", "q4_link_0",
  "q5_other_text", "itr_value", "itr_page",
  "itr_table_num", "itr_supp_link", "general_notes",
  "q1_yes", "q1_no", "q2_yes",
  "q2_no", "q_readme_yes", "q_readme_no",
  "q_readme_unsure", "q3_ps", "q3_ci",
  "q3_un", "q5_r", "q5_python",
  "q5_matlab", "q5_stata", "q5_julia",
  "q5_spss", "q5_mplus", "q5_c_cpp",
  "q5_shell", "q5_unclear", "q5_other",
  "q6_not_mentioned", "q6_broken", "q6_empty",
  "q6_request", "q6_restricted", "q6_other",
  "q_itr_main_yes", "q_itr_main_no", "q_itr_supp_yes",
  "q_itr_supp_no"
)

# Agreement variable per variable to compare
for (v in var_compare) {
  cols <- paste0(v, c("_c1", "_c3", "_c4"))
  combined[[paste0(v, "_agreement")]] <- apply(combined[cols], 1, function(x) length(unique(x)) == 1)
}

# reorder: id, then var1_c1/c2/c3/agreement, var2_c1/c2/c3/agreement, ...
ordered_cols <- c("paper_id", unlist(lapply(var_compare, function(v) 
  c(paste0(v, c("_c1", "_c3", "_c4")), paste0(v, "_agreement"))
)))

combined <- combined %>% select(all_of(ordered_cols))


```

## Comparing variables

```{r}

var_compare <- c("q4_source_0", "q4_link_0",
  "q5_other_text", "itr_value", "itr_page",
  "itr_table_num", "itr_supp_link", "general_notes",
  "q1_yes", "q1_no", "q2_yes",
  "q2_no", "q_readme_yes", "q_readme_no",
  "q_readme_unsure", "q3_ps", "q3_ci",
  "q3_un", "q5_r", "q5_python",
  "q5_matlab", "q5_stata", "q5_julia",
  "q5_spss", "q5_mplus", "q5_c_cpp",
  "q5_shell", "q5_unclear", "q5_other",
  "q6_not_mentioned", "q6_broken", "q6_empty",
  "q6_request", "q6_restricted", "q6_other",
  "q_itr_main_yes", "q_itr_main_no", "q_itr_supp_yes",
  "q_itr_supp_no"
)

```
